\begin{abstract}
\CatchFileBetweenTags{\AlphaInvVal}{calculations/constants.tex}{AlphaInvVal}
\CatchFileBetweenTags{\AlphaSVal}{calculations/constants.tex}{AlphaSVal}
\CatchFileBetweenTags{\WeakAngleVal}{calculations/constants.tex}{WeakAngleVal}

\CatchFileBetweenTags{\HiggsVEVVal}{calculations/constants.tex}{HiggsVEVVal}
\CatchFileBetweenTags{\FermiConstVal}{calculations/constants.tex}{FermiConstVal}
\CatchFileBetweenTags{\HiggsMassVal}{calculations/constants.tex}{HiggsMassVal}
\CatchFileBetweenTags{\ElectronYukawaVal}{calculations/constants.tex}{ElectronYukawaVal}

\CatchFileBetweenTags{\JarlskogEq}{calculations/constants.tex}{JarlskogEq}
\CatchFileBetweenTags{\JarlskogVal}{calculations/constants.tex}{JarlskogVal}
\CatchFileBetweenTags{\CabibboAngleVal}{calculations/constants.tex}{CabibboAngleVal}

\CatchFileBetweenTags{\WBosonMassVal}{calculations/constants.tex}{WBosonMassVal}
\CatchFileBetweenTags{\CabibboAngleEq}{calculations/constants.tex}{CabibboAngleEq}


We derive the fundamental constants of the Standard Model from five geometric integers. By treating the vacuum as a finite-capacity lattice governed by a Persistence Principle (the minimization of Entropic Action), we identify the unique projection of an $E_8$ lattice onto a 4-dimensional manifold. The Standard Model emerges as the Gibbs State of this lattice, the unique thermodynamic ground state satisfying unitarity, causality, and stability constraints.

This projection yields five geometric invariants $\mathbb{S} = \{D{=}4, \Delta{=}43, \nu{=}16, \sigma{=}5, \chi{=}2\}$ that uniquely determine the gauge group ($SU(3) \times SU(2) \times U(1)$), generation number ($n_{\text{gen}} = \sigma - \chi = 3$), and hypercharge assignments, while rendering a fourth generation structurally impossible.

The Standard Model Lagrangian emerges as the Entropic Action of the substrate. We recover the Einstein-Hilbert, Yang-Mills, Dirac, and Higgs terms without modification. Gravity emerges as the Goldstone mode of broken channel-capacity symmetry, governed by an Effective Field Architecture that regulates the lattice flux.

We derive the entire bosonic/structural sector of the Standard Model via global impedance matching: the Fine-Structure Constant ($\alpha^{-1}_{\text{geo}} = \AlphaInvVal \dots$, within $0.6\sigma$ of CODATA 2022); Strong Coupling ($\alpha_s = \AlphaSVal \dots$); Weak Mixing Angle ($\sin^2\theta_W = \WeakAngleVal \dots$); W-Boson Mass ($M_W = \WBosonMassVal$, resolving the CDF/Standard Model tension); Jarlskog Invariant ($J = \JarlskogVal \dots$); Cabibbo Angle ($\theta_C = \CabibboAngleVal$); QCD beta function coefficients ($11 = D\chi + (\sigma - \chi)$, $2/3 = \chi/n_{\text{gen}}$); the complete Higgs sector ($v = \HiggsVEVVal \dots$ GeV, $m_H = \HiggsMassVal \dots$ GeV, $\lambda = 3/23$, $y_e=\ElectronYukawaVal$); and the gravitational hierarchy ($M_P/m_e \propto \alpha^{-\Delta/4}$), resolving the hierarchy problem as geometric attenuation across lattice depth rather than fine-tuning.

This framework replaces parameter fitting with geometric derivation. Falsifiable predictions include: structural prohibition of supersymmetry ($\nu = 16$ saturated), Kaluza-Klein modes ($D = 4$ required), grand unification ($\sigma \neq \chi$), and a zero-degree-of-freedom electroweak fit with $\alpha^{-1}$, $G_F$, and $\sin^2\theta_W$ simultaneously fixed geometrically.

Numerical validation confirms the emergent gravity mechanism ($\kappa = 1.000 \pm 0.001$ across six momentum modes) and a zero-degree-of-freedom global fit of five fundamental observables ($\alpha^{-1}$, $G_F$, $m_W$, $\alpha_s$, $m_H$) yields $\Delta\chi^2 = 1.96$, well below the $3\sigma$ threshold (9.0) confirming a geometric connection.

This is Paper I of a series deriving the fermion spectrum (II), flavor mixing (III), cosmology (IV), and quantum foundations (V).
\end{abstract}