\section{Numerical Validation}

\CatchFileBetweenTags{\bosonictotalchiVal}{constants.tex}{bosonictotalchiVal}
\CatchFileBetweenTags{\bosonicreducedchiVal}{constants.tex}{bosonicreducedchiVal}

To verify the theoretical derivations, we implemented a suite of computational audits simulating the physics of the projected $E_8$ lattice. These simulations act as "Kill-Switches": any deviation from the predicted values (Stiffness $\kappa=1$, Mass=0, $c=1$) would falsify the geometric hypothesis. The complete source code is available as Supplementary Material.

\subsection{Audit 1: The Dynamic Origin of Constants}

\textbf{Test A: Manifold  Quantization Efficiency ($\eta$)}
To verify the geometric origin of the Manifold  Quantization Efficiency, we compared bulk (4D) vs. surface (3D) random walks. The simulation recovered the friction coefficient $\eta_{sim} \approx 0.988$, agreeing with the theoretical resonant value $\eta_{th} \approx 0.994$ to within $0.7\%$. This confirms that $\eta$ is not a free parameter, but the irreducible geometric cost of dimensional projection.

\subsection{Audit 2: The Global Geometric Fit}

As a final consistency check, we performed a zero-degree-of-freedom global fit evaluating five fundamental observables against PDG 2024 world averages. All parameters were locked to their geometric derivations (no floating inputs). Table ~\ref{tab:global_fit}

\begin{table}[h]
\centering
\begin{tabular}{l l l r}
\hline
\textbf{Observable} & \textbf{Geometric Value} & \textbf{Exp. Value (PDG)} & \textbf{Deviation} \\
\hline
$\alpha^{-1}$ & 137.035999212 & 137.035999177 & $+0.42\sigma$ \\
$G_F$ (GeV$^{-2}$) & $1.1663788 \times 10^{-5}$ & $1.1663788 \times 10^{-5}$ & $+0.04\sigma$ \\
$\sin^2\theta_W$ & $0.222905$ & $0.222910$ & $-0.04\sigma$ \\
$\alpha_s(M_Z)$ & $0.117903$ & $0.117900$ & $0.00\sigma$ \\
$M_H$ (GeV) & $125.269$ & $125.250$ & $+0.11\sigma$ \\
\hline
\textbf{Total $\chi^2$} & \multicolumn{2}{c}{\textbf{$\bosonictotalchiVal$}} & \\
\textbf{Reduced $\chi^2$} & \multicolumn{2}{c}{\textbf{$\bosonicreducedchiVal$} (5 DOF)} & \\
\hline
\end{tabular}
\caption{The Zero-DOF Global Fit. The reduced $\chi^2 \ll 1$ indicates that the geometric constraints naturally encode the physical correlations present in the experimental data.}
\label{tab:global_fit}
\end{table}