\section{Synthesis: Structural Audit of the Architecture} \label{sec:Synthesis}

We have derived the fundamental constants of the Standard Model as geometric outputs. However, for this architecture to exist stably, it must satisfy a higher-order constraint: it must minimize Entropic Action ($\delta S_\Phi = 0$) globally. For a distributed regulatory system, this requires \textbf{Global Impedance Matching}—the internal impedances of the regulators must exactly match the geometric apertures of the partitions to prevent information reflection (dissipation).

We conclude this section by verifying that the derived architecture satisfies this condition at three distinct levels of the vacuum structure.

\subsection{Proof 1: Bandwidth Conservation (The Chiral Limit)}
The total allocated capacity must exactly match the hardware limit ($\nu = 16$) without overflow. By summing the derived loads for the Bulk (Gravity), the Boundary (Topology), and the Gauge (Interaction), we verify the conservation of information:

\begin{equation}
\underbrace{B_{res}}_{\text{Gravity}} + \underbrace{\frac{\chi}{\sigma-\chi}}_{\text{Boundary Storage}} + \underbrace{\alpha}_{\text{Gauge}} \stackrel{?}{=} \nu
\end{equation}

Substituting the derived geometric values:
\begin{equation}
15.326... + 0.666... + 0.0073... = 15.9993 \approx 16 \quad (\text{Error} < 0.005\%)
\end{equation}

This precise summation confirms that Gravity is not an external force added to the theory, but the \textbf{inevitable consequence} of using the remaining bandwidth. In a $\nu=16$ universe, the gravitational coupling $\alpha_G$ is strictly determined by the residual bits left after determining the gauge and boundary sectors.

\subsection{Proof 2: Surface Impedance Match (Higgs Subsystem)}
The Surface Regulator (Higgs) must structurally match the aperture of the force it mediates (The Weak Interaction, $\sigma + 1 = 6$). Using the derived Higgs Impedance $Z_H$:

\begin{equation}
Z_H = \frac{1}{\lambda} e^{-2\lambda} \approx 5.97 \approx (\sigma + 1) \quad (\text{Error } < 0.6\%)
\end{equation}

This confirms the local architecture minimizes entropic drag by perfectly matching the vacuum resonance to the weak channel capacity.

\subsection{Proof 3: Core Impedance Match (Gravity Subsystem)}
The Bulk Regulator (Gravity) must achieve unity coupling at the fundamental lattice threshold. Using the derived Planck Mass ($M_P$) and Attenuation Factor ($\alpha^{\Delta/2}$):

\begin{equation}
Z_G(M_P) = \sqrt{\frac{B_{res}}{\alpha^{\Delta/2}}} \cdot \frac{m_e}{M_P} \equiv 1
\end{equation}

This confirms the bulk architecture perfectly compensates for lattice attenuation at the Planck scale.

\subsection{Final Verdict: Global Solvency}
The Effective Field Architecture is globally solvent. The dimensionless ratios (Tier 1) and energy scales (Tier 2) are not independent parameters; they are the unique solutions to the global impedance matching problem. The system persists because it transfers information from the lattice core ($r = \Delta/2$) to the surface ($r = 0$) with minimal dissipative loss, providing the stable stage for the formation of matter.